---
title: "Machine Learning Graphing"
author: "Heather Deel"
date: "2023-09-18"
output: html_document
---

### Libraries
```{r}
library(ggplot2)
library(readxl)
library(ggpubr)
library(paletteer)
library(tidyverse)
library(sjPlot)
library(cowplot)
library(janitor)
library(phyloseq)
library(M3C)
library(vegan)
```

### Comparison R^2 of different data types
```{r}
# tax data was pulled from dan_out/All_stats_randomForest and added to All_stats.csv
ml_stats <- read.csv("machine_learning/16S_EC/All_stats.csv")

p1 <- ggplot(ml_stats, aes(x=myVar, y=r2_val, fill = RA_total_GIBBs)) + 
  geom_boxplot() + 
  xlab("Indicator") +
  ylab(bquote(R^2~"Values")) +
  theme_bw() + 
  ylim(0,1) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_discrete(name = "Data Type", labels = c("EC RA", "EC Total", "EC Modules", "GIBBs RA",
                                                     "GIBBs Total", "Tax RA", "Tax Total")) +
  scale_x_discrete(labels = c("ACE","Active C","Agg. Stab.","Fe","K","Mg","Mn","P",
                              "pH","Respiration","SOM","Water Capacity","Zn"))
p1

ggsave("figures/ml_ec_r2_all.pdf", unit = "in", width = 10, height = 5, dpi = 300, device = "pdf")
ggsave("figures/ml_ec_r2_all.tiff", unit = "in", width = 10, height = 5, dpi = 300, device = "tiff")
```

### Boxplot of R^2 values for each indicator
# individual RA and module graphs
```{r}
ml_stats_RA <- filter(ml_stats, RA_total_GIBBs == "RA")

p1 <- ggplot(ml_stats_RA, aes(x=myVar, y=r2_val)) + 
  geom_boxplot(fill = "#F8766D") + 
  xlab("Indicator") +
  ylab(bquote(R^2~"Values")) +
  theme_bw() + 
  ylim(0,1) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_x_discrete(labels = c("ACE","Active C","Agg. Stab.","Fe","K","Mg","Mn","P",
                              "pH","Respiration","SOM","Water Capacity","Zn"))
p1

#ggsave("figures/ml_ec_r2_RA.pdf", unit = "in", width = 10, height = 5, dpi = 300, device = "pdf")
#ggsave("figures/ml_ec_r2_RA.tiff", unit = "in", width = 10, height = 5, dpi = 300, device = "tiff")

ml_stats_modules <- filter(ml_stats, RA_total_GIBBs == "RA_modules")

p1 <- ggplot(ml_stats_modules, aes(x=myVar, y=r2_val)) + 
  geom_boxplot(fill = "#00BF7D") + 
  xlab("Indicator") +
  ylab(bquote(R^2~"Values")) +
  theme_bw() + 
  ylim(0,1) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_x_discrete(labels = c("ACE","Active C","Agg. Stab.","Fe","K","Mg","Mn","P",
                              "pH","Respiration","SOM","Water Capacity","Zn"))
p1

#ggsave("figures/ml_ec_r2_RA_modules.pdf", unit = "in", width = 10, height = 5, dpi = 300, device = "pdf")
#ggsave("figures/ml_ec_r2_RA_modules.tiff", unit = "in", width = 10, height = 5, dpi = 300, device = "tiff")
```

### Feature importance distributions
```{r}
# coloring by network modules with a correlation coefficient of 0.8
gibbs <- read.csv('picrust2_files/GIBBs.EC.numbers.csv')
gibbs <- gibbs %>%
  filter(Missing != 'x')
gibbs$EC <- gsub(":","_", gibbs$EC)
gibbs$EC <- gsub("\\.","_", gibbs$EC)

myVars <- c('ace', 'SOM', 'activeC', 'resp', 'agg_stab', 'water_cap', 'ph', 'p', 'k', 'mg', 'fe', 'mn', 'zn')

modules <- read.csv("scripts/16S_EC/res_node_table_0.8.csv")
colnames(modules)[1] <- "Predictor"

# models with all enzymes
for (myVar in myVars) {
    filenames <- list.files(paste0("machine_learning/16S_EC/",myVar,"_model_results_RA_corr/"), pattern = paste0(myVar,"_RA_corr_varimp*"), full.names = TRUE)
  varimps <- lapply(filenames, read.csv, header = FALSE)
  varimps_full <- varimps %>% 
    reduce(full_join)

  names(varimps_full) <- c('Predictor', 'VIP', 'Run')
  varimps_full$Predictor <- gsub("EC_","EC:", varimps_full$Predictor)
  varimps_full$Predictor <- gsub("_","\\.", varimps_full$Predictor)

  gDF <- varimps_full %>%
    left_join(gibbs, by=c("Predictor"="EC")) %>% 
    left_join(modules, by = join_by(Predictor)) %>% 
    mutate(Predictor=fct_reorder(Predictor, VIP, .fun=median, .desc=TRUE))
  
  gDF$module <- as.factor(gDF$module)

  # filter out poor indicators
  temp <- gDF %>% group_by(Predictor) %>%
    summarise(Var = median(VIP)) %>%
    top_n(20, Var) 
  
  gDF <- gDF %>%
    filter(Predictor %in% temp$Predictor)
  
  colors <- paletteer_d(`"ggsci::category20_d3"`)
  # colors <- c("Biocontrol" = "#1F77B4FF", 
  #             "C cycling" = "#FF7F0EFF",
  #             "C/N cycling" = "#2CA02CFF",
  #             "N cycling" = "#D62728FF", 
  #             "P cycling" = "#9467BDFF", 
  #             "S cycling" = "#8C564BFF",
  #             "Stess" = "#E377C2FF",
  #             "Siderophore" = "#7F7F7FFF",
  #             "Other" = "white")
  
  means <- aggregate(VIP ~ Predictor, gDF, median)
  means$Category <- 'Other'
  means$VIP <- round(means$VIP, 3)
  p <- ggplot(gDF, aes(x=Predictor, y=VIP, fill=module)) +
    scale_fill_manual(values=colors, na.value="white") +
    geom_boxplot() + 
    geom_point(aes(fill=module), shape=21) + 
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    labs(x="", y="Variable Importance")
  p
    #ggsave(paste0("figures/varimps/", myVar, "_RA_corr_varimp_mod0.8.png"), device='png', dpi=600, height=6, width=8)
}

# models with GIBBs enzymes
for (myVar in myVars) {
     filenames <- list.files(paste0("machine_learning/16S_EC/",myVar,"_model_results_RA_corr_GIBBs/"), pattern = paste0(myVar,"_RA_corr_GIBBs_varimp*"), full.names = TRUE)
  varimps <- lapply(filenames, read.csv, header = FALSE)
  varimps_full <- varimps %>% 
    reduce(full_join)

  names(varimps_full) <- c('Predictor', 'VIP', 'Run')
  varimps_full$Predictor <- gsub("EC_","EC:", varimps_full$Predictor)
  varimps_full$Predictor <- gsub("_","\\.", varimps_full$Predictor)

 gDF <- varimps_full %>%
    left_join(gibbs, by=c("Predictor"="EC")) %>% 
    left_join(modules, by = join_by(Predictor)) %>% 
    mutate(Predictor=fct_reorder(Predictor, VIP, .fun=median, .desc=TRUE))
  
  gDF$module <- as.factor(gDF$module)

  # filter out poor indicators
  temp <- gDF %>% group_by(Predictor) %>%
    summarise(Var = median(VIP)) %>%
    top_n(20, Var) 
  
  gDF <- gDF %>%
    filter(Predictor %in% temp$Predictor)
  
  colors <- paletteer_d(`"ggsci::category20_d3"`)
  # colors <- c("Biocontrol" = "#1F77B4FF", 
  #             "C cycling" = "#FF7F0EFF",
  #             "C/N cycling" = "#2CA02CFF",
  #             "N cycling" = "#D62728FF", 
  #             "P cycling" = "#9467BDFF", 
  #             "S cycling" = "#8C564BFF",
  #             "Stess" = "#E377C2FF",
  #             "Siderophore" = "#7F7F7FFF",
  #             "Other" = "white")
  
  means <- aggregate(VIP ~ Predictor, gDF, median)
  means$Category <- 'Other'
  means$VIP <- round(means$VIP, 3)
  p <- ggplot(gDF, aes(x=Predictor, y=VIP, fill=module)) +
    scale_fill_manual(values=colors, na.value="white") +
    geom_boxplot() + 
    geom_point(aes(fill=module), shape=21) + 
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    labs(x="", y="Variable Importance")
  p
    #ggsave(paste0("figures/varimps/", myVar, "_RA_corr_GIBBs_varimp_mod0.8.png"), device='png', dpi=600, height=6, width=8)
}
```

### Feature importance distributions of models run using 0.8 modules
```{r}
myVars <- c('ace', 'SOM', 'activeC', 'resp', 'agg_stab', 'water_cap', 'ph', 'p', 'k', 'mg', 'fe', 'mn', 'zn')

for (myVar in myVars) {
    filenames <- list.files(paste0("machine_learning/16S_EC/",myVar,"_model_results_modules/"), pattern = paste0(myVar,"_modules_varimp*"), full.names = TRUE)
  varimps <- lapply(filenames, read.csv, header = FALSE)
  varimps_full <- varimps %>% 
    reduce(full_join)

  names(varimps_full) <- c('Predictor', 'VIP', 'Run')

  # filter out poor indicators
  temp <- varimps_full %>% group_by(Predictor) %>%
    summarise(Var = median(VIP)) %>%
    top_n(20, Var) 
  
  gDF <- varimps_full %>%
    filter(Predictor %in% temp$Predictor) %>% 
    mutate(Predictor=fct_reorder(Predictor, VIP, .fun=median, .desc=TRUE))
  
  colors <- paletteer_d(`"ggsci::category20_d3"`)
  
  means <- aggregate(VIP ~ Predictor, gDF, median)
  means$Category <- 'Other'
  means$VIP <- round(means$VIP, 3)
  p <- ggplot(gDF, aes(x=Predictor, y=VIP)) +
    geom_boxplot() + 
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) +
    labs(x="", y="Module Importance")
  p
    ggsave(paste0("figures/varimps/", myVar, "_module_varimp.png"), device='png', dpi=600, height=6, width=8)
}
```

### Relative abundance of ECs by subclasses and subsubclasses
```{r}
ECs <- read.csv("machine_learning/EC_RA_corr/ml_EC_RA_corr.csv", check.names = FALSE)

ECs2 <- ECs[,4:2439]
ECs2_t <- t(ECs2)
ECs2_t <- data.frame(ECs2_t)
ECs2_t <- rownames_to_column(ECs2_t, var = "ECs")

# collapse EC list to subclasses
pattern <- "EC:(.*)\\.(.*)\\."
ECs2_t$ECs_subclass <- regmatches(ECs2_t$ECs, regexec(pattern, ECs2_t$ECs))
ECs2_t$ECs_subclass2 <- sapply(ECs2_t$ECs_subclass, "[[", 2)
ECs2_sc <- ECs2_t[,c(2:537,539)]
ECs2_sc <- ECs2_sc %>% 
  group_by(ECs_subclass2) %>% 
  summarise_all(., .funs = sum)
  
# add EC: back as prefix
ECs2_sc$ECs_subclass2 <- paste0("EC:", ECs2_sc$ECs_subclass2)

# format "otu" table for importing into phyloseq
ECs2_sc <- column_to_rownames(ECs2_sc, var = "ECs_subclass2")

# attach sample names to column names
sample_names <- ECs$SampleID
colnames(ECs2_sc) <- sample_names

OTU = otu_table(ECs2_sc, taxa_are_rows = TRUE)

# format sample data for importing into phyloseq
EC_sam <- ECs[,c(2:3,2440:2588)]
EC_sam <- column_to_rownames(EC_sam, var = "SampleID")
EC_sam$SHMI2_bin <- factor(EC_sam$SHMI2_bin, levels = c("very low","low","medium","high","very high"))
EC_sam$SH_bin <- factor(EC_sam$SH_bin, levels = c("very low","low","medium","high","very high"))

SAM <- sample_data(EC_sam)

# use the EC description as "taxa"
EC_desc <- read_xlsx("ec_files_expasy/enzclass_R.xlsx")
EC_desc <- column_to_rownames(EC_desc, var = "EC")
EC_desc <- as.matrix(EC_desc)
TAX <- tax_table(EC_desc)

# create phyloseq object
p_subclass <- phyloseq(OTU, TAX, SAM)

# make SHMI bar chart
p_df_subclass <- psmelt(p_subclass)
ggplot(p_df_subclass, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "SHMI Bin") +
  scale_fill_discrete(name = "Subclass")
#ggsave("figures/EC_abund_SHMI_subclass.pdf", unit = "in", width = 6, height = 4, dpi = 300, device = "pdf")

# make SEMWISE bar chart
ggplot(p_df_subclass, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "SEMWISE Bin") +
  scale_fill_discrete(name = "Subclass")
#ggsave("figures/EC_abund_SEMWISE_subclass.pdf", unit = "in", width = 6, height = 4, dpi = 300, device = "pdf")

# graph each subclass on its own
# in p_df, create a column with just subclass and filter by that
p_df_subclass$subclass = substr(p_df_subclass$OTU, start = 4, stop = 4)
unique(p_df_subclass$subclass)

# create df for each subclass
# 1 = oxidoreductases
# 2 = transferases
# 3 = hydrolases
# 4 = lyases
# 5 = isomerases
# 6 = ligases

p_df_subclass_1 <- subset(p_df_subclass, subclass == "1")
p_df_subclass_2 <- subset(p_df_subclass, subclass == "2")
p_df_subclass_3 <- subset(p_df_subclass, subclass == "3")
p_df_subclass_4 <- subset(p_df_subclass, subclass == "4")
p_df_subclass_5 <- subset(p_df_subclass, subclass == "5")
p_df_subclass_6 <- subset(p_df_subclass, subclass == "6")

# SHMI bar charts for all levels, combine and save
p1 <- ggplot(p_df_subclass_1, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Oxidoreductases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p2 <- ggplot(p_df_subclass_2, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Transferases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p3 <- ggplot(p_df_subclass_3, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Hydrolases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p4 <- ggplot(p_df_subclass_4, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Lyases") +
  scale_fill_discrete(name = "Subclass") + 
  ylim(0,250)
p5 <- ggplot(p_df_subclass_5, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Isomerases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p6 <- ggplot(p_df_subclass_6, aes(x = SHMI2_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Ligases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p_SHMI_all <- ggarrange(p1,p2,p3,p4,p5,p6,
                    nrow = 2, ncol = 3)
p_SHMI_all <- annotate_figure(p_SHMI_all, 
                              bottom = text_grob("SHMI Bin"),
                              left = text_grob("Abundance", rot = 90))
p_SHMI_all
#ggsave("figures/EC_abund_SHMI_subclass_split.pdf", unit = "in", width = 12, height = 8, dpi = 300, device = "pdf")

# SEMWISE bar charts for all levels, combine and save
p1 <- ggplot(p_df_subclass_1, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Oxidoreductases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p2 <- ggplot(p_df_subclass_2, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Transferases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p3 <- ggplot(p_df_subclass_3, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Hydrolases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p4 <- ggplot(p_df_subclass_4, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Lyases") +
  scale_fill_discrete(name = "Subclass") + 
  ylim(0,250)
p5 <- ggplot(p_df_subclass_5, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Isomerases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p6 <- ggplot(p_df_subclass_6, aes(x = SH_bin, y = Abundance, fill = Family)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Ligases") +
  scale_fill_discrete(name = "Subclass") +
  ylim(0,250)
p_SEMWISE_all <- ggarrange(p1,p2,p3,p4,p5,p6,
                    nrow = 2, ncol = 3)
p_SEMWISE_all <- annotate_figure(p_SEMWISE_all, 
                              bottom = text_grob("SEMWISE Bin"),
                              left = text_grob("Abundance", rot = 90))
p_SEMWISE_all
#ggsave("figures/EC_abund_SEMWISE_subclass_split.pdf", unit = "in", width = 12, height = 8, dpi = 300, device = "pdf")

# subsubclass level
# collapse EC list to subsubclasses
pattern <- "EC:(.*)\\.(.*)\\.(.*)\\."
ECs2_t$ECs_subsubclass <- regmatches(ECs2_t$ECs, regexec(pattern, ECs2_t$ECs))
ECs2_t$ECs_subsubclass2 <- sapply(ECs2_t$ECs_subsubclass, "[[", 1)
ECs2_ssc <- ECs2_t[,c(2:537,541)]
ECs2_ssc$ECs_subsubclass2 = substr(ECs2_ssc$ECs_subsubclass2, 1, nchar(ECs2_ssc$ECs_subsubclass2)-1)
ECs2_ssc <- ECs2_ssc %>% 
  group_by(ECs_subsubclass2) %>% 
  summarise_all(., .funs = sum)

# format "otu" table for importing into phyloseq
ECs2_ssc <- column_to_rownames(ECs2_ssc, var = "ECs_subsubclass2")

# attach sample names to column names
sample_names <- ECs$SampleID
colnames(ECs2_ssc) <- sample_names

OTU = otu_table(ECs2_ssc, taxa_are_rows = TRUE)

# format sample data for importing into phyloseq
EC_sam <- ECs[,c(2:3,2440:2588)]
EC_sam <- column_to_rownames(EC_sam, var = "SampleID")
EC_sam$SHMI2_bin <- factor(EC_sam$SHMI2_bin, levels = c("very low","low","medium","high","very high"))
EC_sam$SH_bin <- factor(EC_sam$SH_bin, levels = c("very low","low","medium","high","very high"))

SAM <- sample_data(EC_sam)

# use the EC description as "taxa"
EC_desc <- read_xlsx("ec_files_expasy/enzclass_R.xlsx")
EC_desc <- column_to_rownames(EC_desc, var = "EC")
EC_desc <- as.matrix(EC_desc)
TAX <- tax_table(EC_desc)

# create phyloseq object
p_subsubclass <- phyloseq(OTU, TAX, SAM)

# make SHMI bar chart
p_df_subsubclass <- psmelt(p_subsubclass)
ggplot(p_df_subsubclass, aes(x = SHMI2_bin, y = Abundance, fill = Genus)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "SHMI Bin") +
  scale_fill_discrete(name = "Sub-subclass")
#ggsave("figures/EC_abund_SHMI_subsubclass.pdf", unit = "in", width = 6, height = 4, dpi = 300, device = "pdf")

# make SEMWISE bar chart
ggplot(p_df_subsubclass, aes(x = SH_bin, y = Abundance, fill = Genus)) +
  geom_bar(stat = "summary", fun = "mean") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "SEMWISE Bin") +
  scale_fill_discrete(name = "Sub-subclass")
#ggsave("figures/EC_abund_SEMWISE_subsubclass.pdf", unit = "in", width = 6, height = 4, dpi = 300, device = "pdf")
```

### Graph relative abundance of modules
```{r}
# create new TAX file, need full EC names
EC_full <- read_csv("ec_files_expasy/enzyme_R.csv")
EC_full <- column_to_rownames(EC_full, var = "Predictor")
EC_full <- as.matrix(EC_full)
TAX <- tax_table(EC_full)

# create sample file with just sample name, SHMI, SEMWISE, for merging with mod files later
EC_sam_SHMI_SEM <- EC_sam
EC_sam_SHMI_SEM <- rownames_to_column(EC_sam_SHMI_SEM, var = "Sample")
EC_sam_SHMI_SEM <- EC_sam_SHMI_SEM %>% 
  select(Sample, SH_rating, SHMI2_rating, SHMI2_bin, SH_bin)

# start with ECs2_t, filter out extra columns
ECs2_t2 <- ECs2_t[,1:537]
ECs2_t2 <- column_to_rownames(ECs2_t2, var = "ECs")

# make column names the sample names
sample_names <- ECs$SampleID
colnames(ECs2_t2) <- sample_names

# make ECs a column for joining
ECs2_t2 <- rownames_to_column(ECs2_t2, var = "Predictor")

# join modules and ECs
ECs_mod <- ECs2_t2 %>% 
  left_join(modules, by = join_by(Predictor))

ECs_mod <- column_to_rownames(ECs_mod, var = "Predictor")
ECs_mod <- ECs_mod[,c(1:536,543)]
ECs_mod <- ECs_mod %>% 
  filter(module != "NA")

# replace the M in module names so I can convert the column to numeric
ECs_mod$module <- gsub("M","", ECs_mod$module)
ECs_mod$module <- as.numeric(ECs_mod$module)

# create module 1 file, phyloseq, and df
ECs_mod8 <- ECs_mod %>% 
  filter(module == 8)
ECs_mod8 <- ECs_mod8[,1:536] # get rid of module col from otu table
OTU8 = otu_table(ECs_mod8, taxa_are_rows = TRUE)
p_mod8 <- phyloseq(OTU8, TAX, SAM)
p_mod8_melt <- psmelt(p_mod8)
p_mod8_persamp <- p_mod8_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod8_persamp$module <- "8"
p_mod8_persamp <- left_join(p_mod8_persamp, EC_sam_SHMI_SEM, by = "Sample")

##### quick graph of ECs_mod1, want to see if they're monotonic
# maybe do a correlation heatmap of everything to verify
library(ggcorrplot)
p.mat <- cor_pmat(ECs_mod8) 
ggcorrplot(ECs_mod8)
ggsave("figures/corr_heatmap_test_mod8.pdf", unit = "in", width = 30, height = 30, dpi = 300, device = "pdf")

# create module 126 file, phyloseq, and df
ECs_mod126 <- ECs_mod %>% 
  filter(module == 126)
ECs_mod126 <- ECs_mod126[,1:536] # get rid of module col from otu table
OTU126 = otu_table(ECs_mod126, taxa_are_rows = TRUE)
p_mod126 <- phyloseq(OTU126, TAX, SAM)
p_mod126_melt <- psmelt(p_mod126)
p_mod126_persamp <- p_mod126_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod126_persamp$module <- "126"
p_mod126_persamp <- left_join(p_mod126_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 206 file, phyloseq, and df
ECs_mod206 <- ECs_mod %>% 
  filter(module == 206)
ECs_mod206 <- ECs_mod206[,1:536] # get rid of module col from otu table
OTU206 = otu_table(ECs_mod206, taxa_are_rows = TRUE)
p_mod206 <- phyloseq(OTU206, TAX, SAM)
p_mod206_melt <- psmelt(p_mod206)
p_mod206_persamp <- p_mod206_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod206_persamp$module <- "206"
p_mod206_persamp <- left_join(p_mod206_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 23 file, phyloseq, and df
ECs_mod23 <- ECs_mod %>% 
  filter(module == 23)
ECs_mod23 <- ECs_mod23[,1:536] # get rid of module col from otu table
OTU23 = otu_table(ECs_mod23, taxa_are_rows = TRUE)
p_mod23 <- phyloseq(OTU23, TAX, SAM)
p_mod23_melt <- psmelt(p_mod23)
p_mod23_persamp <- p_mod23_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod23_persamp$module <- "23"
p_mod23_persamp <- left_join(p_mod23_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 240 file, phyloseq, and df
ECs_mod240 <- ECs_mod %>% 
  filter(module == 240)
ECs_mod240 <- ECs_mod240[,1:536] # get rid of module col from otu table
OTU240 = otu_table(ECs_mod240, taxa_are_rows = TRUE)
p_mod240 <- phyloseq(OTU240, TAX, SAM)
p_mod240_melt <- psmelt(p_mod240)
p_mod240_persamp <- p_mod240_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod240_persamp$module <- "240"
p_mod240_persamp <- left_join(p_mod240_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 272 file, phyloseq, and df
ECs_mod272 <- ECs_mod %>% 
  filter(module == 272)
ECs_mod272 <- ECs_mod272[,1:536] # get rid of module col from otu table
OTU272 = otu_table(ECs_mod272, taxa_are_rows = TRUE)
p_mod272 <- phyloseq(OTU272, TAX, SAM)
p_mod272_melt <- psmelt(p_mod272)
p_mod272_persamp <- p_mod272_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod272_persamp$module <- "272"
p_mod272_persamp <- left_join(p_mod272_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 37 file, phyloseq, and df
ECs_mod37 <- ECs_mod %>% 
  filter(module == 37)
ECs_mod37 <- ECs_mod37[,1:536] # get rid of module col from otu table
OTU37 = otu_table(ECs_mod37, taxa_are_rows = TRUE)
p_mod37 <- phyloseq(OTU37, TAX, SAM)
p_mod37_melt <- psmelt(p_mod37)
p_mod37_persamp <- p_mod37_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod37_persamp$module <- "37"
p_mod37_persamp <- left_join(p_mod37_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 5 file, phyloseq, and df
ECs_mod5 <- ECs_mod %>% 
  filter(module == 5)
ECs_mod5 <- ECs_mod5[,1:536] # get rid of module col from otu table
OTU5 = otu_table(ECs_mod5, taxa_are_rows = TRUE)
p_mod5 <- phyloseq(OTU5, TAX, SAM)
p_mod5_melt <- psmelt(p_mod5)
p_mod5_persamp <- p_mod5_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod5_persamp$module <- "5"
p_mod5_persamp <- left_join(p_mod5_persamp, EC_sam_SHMI_SEM, by = "Sample")

# create module 72 file, phyloseq, and df
ECs_mod72 <- ECs_mod %>% 
  filter(module == 72)
ECs_mod72 <- ECs_mod72[,1:536] # get rid of module col from otu table
OTU72 = otu_table(ECs_mod72, taxa_are_rows = TRUE)
p_mod72 <- phyloseq(OTU72, TAX, SAM)
p_mod72_melt <- psmelt(p_mod72)
p_mod72_persamp <- p_mod72_melt %>% 
  group_by(Sample) %>% 
  summarize(Abundance = sum(Abundance))
p_mod72_persamp$module <- "72"
p_mod72_persamp <- left_join(p_mod72_persamp, EC_sam_SHMI_SEM, by = "Sample")

# merge all module dfs
ace_EC_mods <- full_join(p_mod1_persamp, p_mod126_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod206_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod23_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod240_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod272_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod37_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod5_persamp)
ace_EC_mods <- full_join(ace_EC_mods, p_mod72_persamp)

# line graph, also playing with boxplot and SHMI2_bin - unsure which is best, but either way with ACE I'm not seeing much of a difference
# may need to put module 1 in it's own graph or something? Not seeing the trends per module like I wanted to
p_mod_ace <- ggplot(ace_EC_mods, aes(x = SHMI2_bin, y = Abundance, color = module)) +
  geom_boxplot()
p_mod_ace

################ going to move away from SHMI - want to run networks with modules in which cooccurring enzymes are trending in the same direction, will then run RF models with those modules to get rid of the collinearity and then we can interpret the modules (what enzymes are in the important modules? are there any in our gibbs list? are there any that are "important" simply because they cooccur with something else in the soil, and that something else is what's actually being driven by management?)
```

### PCOAs using tsne
```{r}
P <- readRDS('machine_learning/EC_RA_corr/ml_EC_RA_corr.RDS')

gibbs <- read.csv('picrust2_files/GIBBs.EC.numbers.csv')
gibbs <- gibbs %>%
  filter(Missing != 'x')

data <- P %>%
  select(gibbs$EC)

meta <- P

#tsne_out <- Rtsne::Rtsne(data, perplexity=15, max_iter=1000)
#plot(tsne_out$Y)

p1 <- M3C::tsne(t(data), labels=FALSE, dotsize = 0, seed=1234, perplex=15) + theme_bw() +
  geom_point(aes(fill=meta$ClimateZ, shape=meta$Current_Land_Use), size=4) +
  scale_shape_manual(values=c(21,22,23,24)) +
  guides(size="none",
         fill = guide_legend(override.aes = list(size = 4, shape=21))) +
  labs(fill="ClimateZ", shape="Land Use", x="tSNE 1", y="tSNE 2")
p1
ggsave(paste0("figures/All_GIBBs_RA_corr_tSNE.png"), device='png', dpi=600, height=4, width=6)
```

### quick plot of climate by activeC
# and do a feature
```{r}
p <- ggplot(ml_EC_16S, aes(x = ClimateZ, y = activeC, fill = ClimateZ)) +
  geom_boxplot()
p

p <- ggplot(ml_EC_16S, aes(x = clay, y = activeC)) +
  geom_point() +
  geom_smooth()
p

p <- ggplot(ml_EC_16S, aes(x = `EC:1.10.9.1`, y = activeC)) +
  geom_point() +
  geom_smooth()
p

p <- ggplot(ml_EC_16S, aes(x = `EC:1.3.99.35`, y = activeC)) +
  geom_point() +
  geom_smooth()
p

p <- ggplot(ml_EC_16S, aes(x = `EC:3.1.1.81`, y = activeC)) +
  geom_point() +
  geom_smooth()
p

```





















