---
title: "18S_EC"
author: "Heather Deel"
date: "2023-09-22"
output: html_document
---

### Setup and packages
```{r setup, include=FALSE}
library(tidymodels)
library(tidyverse)
library(workflows)
library(tune)
library(metagMisc)
library(ranger)
library(randomForest)
```

### Load and format data
# only need to run this once - can skip to the next chunk
# run on local computer
```{r}
# we don't need the full phyloseq from before, so reimporting
func <- read.delim('picrust2_files/18S/EC_metagenome_out/pred_metagenome_contrib.tsv.gz', sep='\t')
df <- read.delim('picrust2_files/18S/picrust2_out_SHAI_18S/marker_nsti_predicted.tsv.gz', sep = '\t')

names(df) <- c("taxon", "genome_16S_count", "metadata_NSTI")

# combine two files
func <- func %>%
  full_join(df, by="taxon")

# get other metadata
data.pred.SHMI <- readRDS('/project/soil_micro_lab/micro_indicators/machine_learning/data.pred.SHMI.RDS')

# read in metadata that includes sampleIDs to merge with data.pred.SHMI
data.sampleID <- read.delim('/project/soil_micro_lab/micro_indicators/machine_learning/SHAI.Meta.22March2023.q2.reduced.txt', sep = '\t')

# merge metadata files
data.merged <- merge(data.pred.SHMI, data.sampleID, by = "PLFA_ID", all.x = TRUE)

# move the sample ID column to the front
data.merged <- data.merged %>% 
  relocate(SampleID)

# get qPCR data into func
# need to read in raw qPCR values
qPCR_data <- read.delim('/project/soil_micro_lab/micro_indicators/machine_learning/SHAI.Meta.22March2023.q2.qPCR.txt', sep = '\t')
qPCR_data <- qPCR_data %>% 
  filter(!is.na(qPCR))

func <- merge(func, qPCR_data, by = "sample")

# we imported un-normalized counts, so normalizing here
func$gene_counts <- func$taxon_rel_abun / 100 * func$genome_function_count / func$genome_16S_count * func$qPCR

# change column name "function" to "EC"
colnames(func)[2] <- "EC"

# pool data by sample and function
otu <- func %>%
  group_by(sample, EC) %>%
  summarize(sum=sum(gene_counts))
otu <- otu %>%
  spread(key=sample, value=sum)
otu <- data.frame(otu)
otu[is.na(otu)] <- 0
row.names(otu) <- otu$EC
otu <- otu[,-1]

# transpose the "otu" table
otu <- t(otu)

# convert matrix to data frame
otu <- as.data.frame(otu)

# make the sample names a column for merging later
otu <- tibble::rownames_to_column(otu, "SampleID")

# merge metadata and "otu" table
hops2018_all <- merge(data.merged, otu, by = "SampleID")
hops2018_all <- as.data.frame(hops2018_all)

# select columns to be used in ml - SampleID, ECs, and "Overall" CASH rating
hops2018_ml <- hops2018_all[,c(1,95,152:2390)]

saveRDS(hops2018_ml, "/project/soil_micro_lab/micro_indicators/machine_learning/16S_EC/hops2018_ml_EC.RDS")
hops2018_ml <- readRDS("/project/soil_micro_lab/micro_indicators/machine_learning/16S_EC/hops2018_ml_EC.RDS")